---
title: "graphs"
output: html_document
---
# libraries
```{r}
library(tidyverse)
library(here)
library(rio)
library(ggplot2)
```

# data
## import and clean dataframes
```{r}
# conservative-motivated
arctic_df <- import(here("data", "arctic-ice-data.csv"))
temp_df <- import(here("data", "global-temp-data.txt"))

# liberal-motivated
ozone_df <- import(here("data", "ozone-data.txt")) %>% 
  janitor::clean_names() %>% 
  rename(
    "year" = "v1",
    "area" = "value"
  )

# boundary conditions
co2_df <- import(here("data", "carbon-dioxide-data.txt"))
co2_df <- co2_df[2:nrow(co2_df),] %>% 
  rename(
    "year" = "V1",
    "monthly_avg" = "V4",
    "deseasonalized" = "V5"
  )

icesheets_df <- import(here("data", "ice-sheets-data.txt")) %>% 
  rename(
    "year" = "V1",
    "mass" = "V2",
    "uncertainty" = "V3"
  )
## fix year value
icesheets_df$year <- as.character(icesheets_df$year) %>% 
  strtrim(., 4)
 
# other
sealevel_df <- import(here("data", "sea-level-data.txt"))
```

## create plot dataframes
```{r}
arcticplot_df <- arctic_df %>% 
  select(year, area)

tempplot_df <- temp_df %>% 
  select(Year, No_Smoothing)

ozoneplot_df <- ozone_df %>% 
  select(year, area)

co2plot_df <- co2_df %>% 
  select(year, monthly_avg, deseasonalized)

icesheetsplot_df <- icesheets_df %>% 
  select(year, mass)
```

## create jittered plot dataframes
```{r}
# j725_arcticplot_df <- arcticplot_df %>% 
#  mutate(area = jitter(area, amount = .725))

jitteredilikeit <- j725_arcticplot_df
```

# plots
## political plots
### arctic plot
```{r}
arcticplot <- arcticplot_df %>% 
  ggplot(aes(year, area)) +
  geom_point() +
  geom_line()

arcticplot + 
  labs(x = NULL,
       y = "Amount of Ice (million square kilometers)",
       title = "Average Monthly Amount of Arctic Sea Ice \n 1979-2013") +
  theme(
    axis.text.x = element_text(face = "bold",
                               color = "black",
                               size = 10,
                               angle = 45),
    axis.text.y = element_text(face = "bold",
                               color = "black",
                               size = 10)
  )
  
```

### temp plot
```{r}
tempplot <- tempplot_df %>% 
  filter(Year >= 1980) %>% 
  ggplot(aes(Year, No_Smoothing)) +
  geom_point() +
  geom_line()
  
```

### ozone plot
```{r}
ozoneplot <- ozoneplot_df %>% 
  filter(year != 2021) %>% 
  ggplot(aes(year, area)) +
  geom_point() +
  geom_line()

ozoneplot <- ozoneplot_df %>% 
  filter(year >=2000 & year != 2021) %>% 
  ggplot(aes(year, area)) +
  geom_point() +
  geom_line()
  
```

### co2 plot --- issue here bc i need to avg the monthly avgs to make an annual avg
```{r}
co2plot <- co2plot_df %>% 
  ggplot(aes(year, monthly_avg)) +
  geom_point() +
  geom_line()
```

### ice sheets -- also issue bc need to avg
```{r}
icesheetsplot <- icesheetsplot_df %>% 
  ggplot(aes(year, mass)) +
  geom_point() +
  geom_line()
```

## non-political plots
### arctic jittered plot
```{r}
j725_arcticplot_df %>% 
  ggplot(aes(year, area)) +
  geom_point() +
  geom_line()

arcticplot_df %>% 
  ggplot(aes(year, area)) +
  geom_point() +
  geom_line()
```

### temp jittered plot
```{r}

```

### ozone jittered plot
```{r}

```

### co2 jittered plot
```{r}

```

### ice sheets jittered plot
```{r}

```

